---
title: "hw1_program"
author: "Subhrajyoty Roy"
date: "August 11, 2019"
output: 
    pdf_document:
        keep_tex: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
```

```{r}
library(vcdExtra)
data("Abortion")   #loads the Abortion data
Abortion
```

Note that, current data is in a format of 2x2 table with *Sex* and *Status* as row and column, while *Support_Abortion* as layer. However, we need to restructure the data in a way so that *Sex* and *Support_Abortion* to be row and column variable respectively, while *Status* is layer variable.

```{r}
Abortion2 = aperm(Abortion, c(1,3,2))
dimnames(Abortion2)
```

Now, to test the null hypothesis $H_0: \text{row}||\text{column}|\text{Layer}$ we use **Cochran-Mantel-Haenszel Chi-Squared Test**;

```{r}
mantelhaen.test(Abortion2)
```

Note that, we get an p-value of $0.004826$, which is extremely lower than the significance level of $\alpha = 0.05$. Hence, we reject the null hypothesis that the variables *Sex* and *Support_Abortion* are independent given the layered variable *Status* in the light of *Abortion* data.







