---
title: "HW2 Program"
author: "Subhrajyoty Roy"
date: "August 25, 2019"
output: 
    pdf_document:
        keep_tex : true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "", warning = FALSE)
```


We first load the required packages in *R*.

```{r}
library(vcd)
library(lpSolve)
```


Now, we load the *SexualFun* data from **vcd** package.

```{r}
data <- vcd::SexualFun
ftable(data)
```


We first obtain the actual values of Cohen's Kappa $\kappa$ and the measure $\lambda_r$ for the actual table;

```{r}
kp <- Kappa(data)   #compute the actual Kappa 
print(kp)
```

We note that the actual value of Cohen's Kappa for *SexualFun* data is $0.12933025$.

Consider the following code which computes $\lambda_r$;

```{r}
n <- sum(data)
rowMar <- rowSums(data)/n
colMar <- colSums(data)/n

a <- sum(diag(data))/n
b <- max((rowMar + colMar)/2)

lambda <- (a - b)/(1-b)
print(lambda)
```

We find that the actual value of $\lambda_r$ for *SexualFun* data is $0$.


Now, we consider the minimization and maximization problem, where we try to find extremal values of $\sum_i p_{ii}$ or correspondingly $\sum_i n_{ii}$. We shall use **lpSolve** package to solve the corresponding linear programming for us.

```{r}
lpMax = lp.transport(cost.mat = diag(4), direction = "max", 
             row.signs = rep("==", 4), row.rhs = rowSums(data), 
             col.signs = rep("==", 4), col.rhs = colSums(data))

```

The maximizing configuration would look like;

```{r}
print(lpMax$solution)
```

Therefore, the maximum value of Cohen's kappa and $\lambda_r$ is obtained using the following code;

```{r}
Kappa(lpMax$solution)

a <- lpMax$objval/n
lambda <- (a - b)/(1-b)
print(lambda)
```

Therefore, the maximum value of Cohen's Kappa given the marginals is $0.8799$, while the maximum value of $\lambda_r$ given the marginals is $0.8621$. 

We use similar method to find the minimum value of these measures given the marginals.

```{r}
lpMin = lp.transport(cost.mat = diag(4), direction = "min", 
             row.signs = rep("==", 4), row.rhs = rowSums(data), 
             col.signs = rep("==", 4), col.rhs = colSums(data))

```

The maximizing configuration would look like;

```{r}
print(lpMin$solution)
```

```{r}
Kappa(lpMin$solution)

a <- lpMin$objval/n
lambda <- (a - b)/(1-b)
print(lambda)
```

Therefore, the maximum value of Cohen's Kappa given the marginals is $-0.3661$, while the maximum value of $\lambda_r$ given the marginals is $-0.5689$. 











